Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites

### 2020-11-30 doc/SA-Hermaphrodites-wDemography.tex
Current edits to ms; Handover Intro to Lotte
Just pushing edits for the current state of the ms, along with minor
changes to the makeDeltaPolySpaceData() function so the output has \
a different filename.

Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}. Such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching the phenotypic optimum in one or more fitness components, thereby increasing the population's overall extinction risk. At the same time, they provide an effective mechanism for the maintenance of genetic variation, which in turn influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}. Hence, for genetically constrained traits the nature and extent of genetic variation observed in natural populations should reflect the balance between the contribution of constraints to the maintainance of genetic polymorphisms, and the demographic consequences of the resulting maladaptation.
%Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
%The nature and extent of genetic variation observed in natural populations should therefore reflect the balance between the contribution of those constraints to the maintainance of genetic polymorphism and the demographic consequences of the resulting maladaptation.

Genetic trade-offs between male and female fitness are a common feature of sexually reproducing populations, and are thought to contribute not only to the maintenance of genetic variation, but also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}. When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}. Moreover, alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.

@@

In this article, we build upon the models of \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b} to develop a stage-structured mendelian matrix model that accommodates selection through male and female function in simultaneous hermaphrodites with mixed mating systems. Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations. Although a variety of interesting eco-evolutionary dynamics are possible in the model (e.g., \citealt{deVriesCaswell2019a, deVriesCaswell2019b,CaswellEtAl2018}), we focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate). The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions. Finally, we give an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).

A key finding from our analysis is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism. Furthermore, this demographically viable polymorphic parameter space is often biased towards 'female-beneficial/male-deleterious' SA alleles (i.e., those with weaker selection on the female- than male sex function). Under some conditions, self-fertilization can alleviate the demographic costs of balanced SA polymorphisms, but the concommitant effects of inbreeding depression greatly exacerbate them. Interestingly, our illustrative example using {\itshape M.~guttatus} appears to show that demographic rates observed in real populations can sustain large regions of viable SA polymorphic space. Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.

@@

Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites. The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing. A full derivation of the model and analyses is presented in \hl{Appendix A}, and all simulation code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.

@@

Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and $\delta$ represents the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}. In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}). Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b}, except each individual's genotype determines both the number of ovules produced, and also paternal "mating success", broadly defined. For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986,WangBarrett2020,Harder2016}. Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the male stage $\times$ genotype vector:
\begin{linenomath*}
\begin{equation}
		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} = 
		\left(
			\begin{array}{ccc}
				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & 0 \\
				0 & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
			\end{array} \right)
		\left(
			\begin{array}{ccc}
				\mbf{F}^{\prime}_{AA} & 0 & 0 \\
				0 & \mbf{F}^{\prime}_{Aa} & 0 \\
				0 & 0 & \mbf{F}^{\prime}_{aa} \\
			\end{array} \right)
		\left(
			\begin{array}{c}
				\mbf{p}_{AA} \\
				\mbf{p}_{Aa} \\
				\mbf{p}_{aa} \\
			\end{array} \right).
\end{equation}
\end{linenomath*}

@@

Reproduction via self-fertilization differs from outcrossing. As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection from external factors relative to outcrossing (e.g., \citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}). Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during meiosis: 
\begin{linenomath*}
\begin{equation} \label{eq:HS}
	\mbf{H}^S(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				1 & 1/4 & 0 \\
				0 & 1/2 & 0 \\
				0 & 1/4 & 1 \\
			\end{array} \right).
\end{equation}
\end{linenomath*}

@@

Using the component matrices described in the previous section, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
\begin{linenomath*}
\begin{equation} \label{eq:Atilde}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}}.
\end{equation}
\end{linenomath*}

@@

We now construct and analyze a genotype $\times$stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions. To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active. Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect. Following convention, we parameterize the relative fitnesses of each genotype through each sex function ($w_{i}$ and $w^{\prime}_{i}$) to be bounded by $(0,1)$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}). The SA locus does not affect survival and transition rates. However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing. By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovlues. With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
\begin{linenomath*}
\begin{equation} \label{eq:US}
	\mbf{U}^S = \left(
					\begin{array}{cc}
						\sigma^S_J (1 - \delta_J) \big(1 - \gamma^S (1 - \delta_{\gamma}) \big) & 0 \\
						\sigma^S_J (1 - \delta_J) \gamma (1 - \delta_{\gamma})      & \sigma^S_A (1 - \delta_S)
					\end{array}
				\right) \\
\end{equation}
\end{linenomath*}
\noindent and 
\begin{linenomath*}
\begin{equation}\label{eq:UX}
	\mbf{U}^X = \left(
					\begin{array}{cc}
						\sigma^X_J(1 - \gamma^X) & 0 \\
						\sigma^X_J \gamma      & \sigma^X_A
					\end{array}
				\right)
\end{equation}
\end{linenomath*}

\noindent where $\sigma_J$ and $\sigma_A$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_S$, $\delta_A$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}). Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility. For simplicity, we assume these are constant among genotypes.

@@

@@

We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium allele frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}). The formal conditions for a protected polymorphism are determined by linearizing the model in the vacinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, which we denote $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$. For our model, $\mbf{M}$ has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}). For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
\begin{linenomath*}
\begin{equation}\label{eq:PolyCond}
	1 < \frac{\rho(\mbf{M}_{22})}{\lambda_{i}}
\end{equation}
\end{linenomath*}

@@

Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for very restricted conditions (see \hl{Online Appendix X}). We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space. Because single-locus selection coefficients are generally weak, we limit our analyses to coefficients with within the open interval $\left(0,0.15\right)$, unless stated otherwise. For a given parameter set, we numerically evaluated Eq(\ref{eq:PolyCond}) and $\lambda$, starting at the outer boundaries of selection parameter space, and titrated inward to identify thresholds where $\tilde{\zeta}_i \approx 1$ and $\lambda \approx 1$ (with precision of $10^{-5}$). The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds

>>>>

Conflicting selection and gene-flow between different classes of individuals, or genetic trade-offs between fitness components, can impose constraints on adaptation with particularly interesting evolutionary consequences \citep{CharlesworthHughes2000, ConnallonHall2018}. On the one hand, such constraints hinder adaptation by preventing individuals (or classes of individuals) from reaching their phenotypic optimum in one or more fitness components, which can increase the population's overall extinction risk. On the other hand, they provide an effective mechanism for the maintenance of genetic variation, which influences the population's capactiy for future adaptation \citep{Fisher1930, CharlesworthHughes2000, ConnallonHall2018, MatthewsConnallon2019}. Hence, for traits under conflicting selection, the nature and extent of genetic variation observed in natural populations should reflect a balance between the contribution of genetic constraints to the maintainance of genetic polymorphisms, and the demographic consequences of the resulting maladaptation.
%Genetic trade-offs between fitness components due to pleiotropy represent a form of genetic constraint with particularly interesting evolutionary consequences because they act as a double-edged sword
%The nature and extent of genetic variation observed in natural populations should therefore reflect the balance between the contribution of those constraints to the maintainance of genetic polymorphism and the demographic consequences of the resulting maladaptation.

Genetic trade-offs between male and female fitness are a common feature of sexually reproducing populations, and are thought to contribute not only to the maintenance of genetic variation, but also the evolution of sexual dimorphism and gender-related traits \citep{Lande1980, Rice1992, Charlesworth1999, RiceChippindale2001, BondurianskyChenoweth2009,Olito2019}. When such genetic trade-offs occur, "sexually antagonistic" selection (abbreviated SA hereafter) can arise when beneficial alleles for one sex are deleterious when expressed in the other \citep{Kidwell1977, Rice1992, ConnallonClark2012}. This, alleles with opposing fitness effects through male and female sex functions create analogous genetic constraints on fitness in hermaphrodite populations, where both maternal and paternal reproductive success contribute jointly to each individuals' overall fitness \citep{LloydWebb1986, WebbLloyd1986, Abbott2011, JordanConnallon2014}.

@@

In this article, we develop a stage-structured mendelian matrix model that accommodates selection through male and female sex-functions in populations of simultaneous hermaphrodites with mixed mating systems. Under obligate outcrossing, the model becomes mathematically identical to the two-sex model of \citet{deVriesCaswell2019b}. Using the model, we jointly analyze the evolutionary and demographic consequences of SA selection in obligately outcrossing (equivalent to dioecious) and partially selfing hermaphrodite populations. We focus our analyses on identifying the parameter conditions under which SA polymorphisms are maintained by balancing selection {\itshape and} the population is also demographically viable (i.e., has a positive intrinsic growth rate). The model also allows for the effects of inbreeding depression to manifest in different life-history stages, enabling us to explore the fitness consequences of early- vs.~late-acting inbreeding depression on the model predictions. Finally, we give an illustrative example of the potential for SA polymorphism in real populations by parameterizing our model with empirically estimated demographic rates for the hermaphroditic flowering plant {\itshape Mimulus guttatus} (now {\itshape Erythranthe guttata}).

A key finding of our study is that when the intrinsic population growth rate is close to $1$, the deleterious effects of segregating male-beneficial SA alleles on female fecundity can result in extinction over much of the parameter space for SA polymorphism. Furthermore, this demographically viable polymorphic parameter space is often biased towards alleles with stronger selection through the female- than male sex function. Self-fertilization can alleviate the demographic costs of balanced SA polymorphisms under some conditions, however, the concommitant effects of inbreeding depression generally exacerbate them in populations with mixed-mating systems. Despite these demographic pitfalls, our example using {\itshape M.~guttatus} appears to show that demographic rates observed in real populations are capable of sustaining large regions of viable SA polymorphic space. Overall, our findings provide a more nuanced picture of the nature of SA genetic variation that we should expect to find in natural populations, where the fate of SA alleles and the populations harboring them is determined jointly by evolutionary and demographic processes.

@@

Here, we briefly describe a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female fertility for populations of simultaneous hermaphrodites with outcrossing or mixed-mating systems. The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing. A full derivation of the model and analyses is presented in \hl{Appendix A}, and all simulation code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.

@@

Let $C$ denote the proportion of each individual's ovules that are self-fertilized (the remaining $1 - C$ are outcrossed), and let $\delta$ represent the proportion of self-fertilized zygotes that fail to develop due to inbreeding depression during early development \citep{Charlesworth1987}. In \hl{Appendix A.x}, we present a more general model of genotype-specific self-fertilization rates (after \citealt{JordanConnallon2014}). Outcrossing proceeds similarly to a two-sex model of reproduction \citep{deVriesCaswell2019b} except that each individual's genotype determines both the number of ovules produced and also paternal "mating success", broadly defined. For hermaphroditic flowering plants, for example, paternal mating success could reflect pollen production, export efficiency, pollen-tube germination and growth rates, among other things \citep{LloydWebb1986, WangBarrett2020, Harder2016}. Depending on their stage and genotype at the SA locus, individuals contribute male gametes to the overall population gamete pool according to the male stage $\times$ genotype vector:
\begin{linenomath*}
\begin{equation}
		\mbf{W}^{\prime} \mbb{F}^{\prime} \mbf{p} = 
		\left(
			\begin{array}{ccc}
				\mbf{1}^{\intercal}_{\omega} & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & 0 \\
				0 & \frac{1}{2} \mbf{1}^{\intercal}_{\omega} & \mbf{1}^{\intercal}_{\omega} \\
			\end{array} \right)
		\left(
			\begin{array}{ccc}
				\mbf{F}^{\prime}_{AA} & 0 & 0 \\
				0 & \mbf{F}^{\prime}_{Aa} & 0 \\
				0 & 0 & \mbf{F}^{\prime}_{aa} \\
			\end{array} \right)
		\left(
			\begin{array}{c}
				\mbf{p}_{AA} \\
				\mbf{p}_{Aa} \\
				\mbf{p}_{aa} \\
			\end{array} \right).
\end{equation}
\end{linenomath*}

@@

Reproduction via self-fertilization differs from outcrossing. As in previous models of partial selfing, we assume that individuals produce enough male gametes to easily self-fertilize all of their ovules, and that self-fertilization involves little or no selection from external factors relative to outcrossing (e.g., \citealt{Charlesworth1978a,JordanConnallon2014,Olito2017}, but see \citealt{Tazzyman2015}). Under these assumptions, the genotype distributions of selfed offspring are determined entirely by the parental genotype and the probabilities of segregation and fertilization during and after meiosis: 
\begin{linenomath*}
\begin{equation} \label{eq:HS}
	\mbf{H}^S(\tilde{\mbf{n}}) = 
			\left(
			\begin{array}{ccc}
				1 & 1/4 & 0 \\
				0 & 1/2 & 0 \\
				0 & 1/4 & 1 \\
			\end{array} \right).
\end{equation}
\end{linenomath*}

@@

Using the component matrices described above, the matrix $\tilde{\mbf{A}}[\tilde{\mbf{n}}]$ that projects the eco-evolutionary dynamics is:
\begin{linenomath*}
\begin{equation} \label{eq:Atilde}
	\tilde{\mbf{A}}[\tilde{\mbf{n}}] = 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{U}^S & \mbf{0} \\ \hline
				\mbf{0} & \mcal{U}^X \\
			\end{array} \right)}_{\tilde{\mbf{U}}} + 
			\underbrace{\left(
			\begin{array}{c|c}
				\mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) & \mcal{F}^S(\tilde{\mbf{p}}) C(1 - \delta) \\ \hline
				\mcal{F}^X(\tilde{\mbf{p}}) (1 - C) & \mcal{F}^X(\tilde{\mbf{p}}) (1 - C)\\
			\end{array} \right)}_{\tilde{\mbf{F}}}.
\end{equation}
\end{linenomath*}

@@

We now construct and analyze a genotype $\times$ stage-classified model for a hypothetical species with intralocus sexual conflict via the two sex functions. To keep the model analytically tractable, we assume our hypothetical species has two life stages: juveniles and adults (i.e., $\omega = 2$), and that only adults are reproductively active. Suppose that there is a genetic trade-off between the sex-functions at a single bi-allelic locus such that allele $A$ is beneficial for female fertility but detrimental for male reproductive success (e.g., pollen production), and that allele $a$ has the reverse effect. Following convention, we parameterize the relative fitnesses of each genotype through each sex function ($w_{i}$ and $w^{\prime}_{i}$) to be bounded by $(0,1)$, with dominance and selection coefficients $h_f, s_f$ and $h_m, s_m$ determining the decrease in fertility through each sex function relative to the most fit genotype ($AA$ has highest female fertility, $aa$ the highest male fertility; see Table \ref{tab:Fitness}). The SA locus does not affect survival and transition rates. However, the survival matrices can be used to model the fitness effects of inbreeding depression at later stages of development by allowing different stage-specific survival and transition rates for individuals produced by self-fertilization vs.~outcrossing. By contrast, the parameter $\sigma$ only affects inbreeding depression through viability of selfed ovlues. With this in mind, we define survival matrices for individuals produced by selfing and outcrossing as follows:
\begin{linenomath*}
\begin{equation} \label{eq:US}
	\mbf{U}^S = \left(
					\begin{array}{cc}
						\sigma_j (1 - \delta_j) \big(1 - \gamma (1 - \delta_{\gamma}) \big) & 0 \\
						\sigma_j (1 - \delta_j) \gamma (1 - \delta_{\gamma})      & \sigma_a (1 - \delta_a)
					\end{array}
				\right) \\
\end{equation}
\end{linenomath*}
\noindent and 
\begin{linenomath*}
\begin{equation}\label{eq:UX}
	\mbf{U}^X = \left(
					\begin{array}{cc}
						\sigma_j(1 - \gamma) & 0 \\
						\sigma_j \gamma      & \sigma_A
					\end{array}
				\right)
\end{equation}
\end{linenomath*}

\noindent where $\sigma_j$ and $\sigma_a$ are the juvenile and adult stage survival rates, $\gamma$ is the maturation rate from juvenile to adult stages, and the corresponding $\delta_j$, $\delta_a$, and $\delta_{\gamma}$ terms denote the proportional decreases in stage-specific survival and transition rates due to inbreeding depression (i.e., deleterious effects of inbreeding at later life-history stages; e.g., \citealt{HarderRoutely2006}). Any or all of these parameters may differ among genotypes, so that selection can operate on stage-specific viability, development, and/or fertility. For simplicity, we assume these are constant among genotypes.

@@

Unless stated otherwise, we use the following parameter values for the demogrphic rates in the model: $\sigma_j = \sigma_A = 0.6$ and $\gamma = 0.05$, which are similar to those used in \citet{deVriesCaswell2019b}. Our parameters of interest include $f$, $\delta$, $\delta_j$, $\delta_a$, $\delta_{\gamma}$, and the selection parameters $h_f$, $s_f$, $h_m$, and $s_m$, which are given different values for each analyses as described in the figure captions.

@@

We identify conditions where SA polymorphism is 'protected' by evaluating the stability of populations initially fixed for either SA allele to invasion by the other (i.e., stability was assessed at the boundary equilibrium genotype frequencies of $\hat{\mbf{p}}_{AA} = 1$ and $\hat{\mbf{p}}_{aa} = 1$; \citealt{Levene1953, Prout1968, deVriesCaswell2019b}). The formal conditions for a protected polymorphism are determined by linearizing the model in the vacinity of the boundary equilibria ($\hat{\mbf{p}}$), and evaluating the magnitude of the largest eigenvalue of the Jacobian matrix of the linearization, which we denote $\tilde{\zeta}_{AA}$ and $\tilde{\zeta}_{aa}$. For our model, the Jacobian, $\mbf{M}$, has an upper block triangular form, and we can therefore express the stability of $\hat{\mbf{p}}$ in terms of the leading eigenvalue of the diagonal block corresponding to changes in heterozygote frequency ($\mbf{M}_{22}$; e.g., see \citealt{OttoDay2007,deVriesCaswell2019b}). For a polymorphism to be 'protected', both boundary equilibria must be unstable (i.e., $\tilde{\zeta}_{AA} > 1$ and $\tilde{\zeta}_{aa} > 1$), which requires that
\begin{linenomath*}
\begin{equation}\label{eq:PolyCond}
	1 < \frac{\rho(\mbf{M}_{22})}{\lambda_{i}}
\end{equation}
\end{linenomath*}

@@

Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for very restricted conditions (see \hl{Online Appendix X}). We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space. Because single-locus selection coefficients are generally weak (i.e., $s < 0.1$; e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise. For a given parameter set, we numerically evaluated Eq(\ref{eq:PolyCond}) and $\lambda$, starting at the outer boundaries of selection parameter space, and titrated inward to identify thresholds where $\tilde{\zeta}_i \approx 1$ and $\lambda \approx 1$ (with precision of $10^{-5}$). The proportions of total $s_f \times s_m$ parameter space where SA polymorphism and/or extinction occurred were calculated by numerically integrating the appropriate regions defined by the invasion and extinction thresholds

%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Demographic rate estimates for {\itshape M. guttatus}} \label{sec:MguttMethods}

To illustrate the use of our model with empirical data, and see whether demographic rates observed in natural populations are likely to support balanced SA polymorphisms, we used publicly available demographic and fitness data for natural populations of the hermaphroditic flowering plant {\itshape Mimulus guttatus} ({\itshape Erythranthe guttata}). {\itshape M.~guttatus}
